---
title: "missBART"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{my_vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo=FALSE}
library(missBART)
```

## Introduction
The package __missBART__ was created for modelling multivariate, non-linear data with non-ignorable missingness in the response. There are multiple functions within the package, such as functions for specifying hyperparameters, functions for building CART trees, functions for sampling from multivariate distributions, and the main function `missBART.probit`, which fits a multivariate BART and probit regression model to a dataset with non-ignorable response values. 

## Sampling functions and tree functions
The sampling functions such as `multi_rMVN()` and `matrnorm()` can be used to simulate data from a multivariate normal or matrix normal distribution. To simulate data from a sum-of-trees model, we can use `propose_tree()` to iteratively build our tree structures.

## Example with simulated data
To demonstrate the use of the different functions in the package, we will simulate a dataset from a sum-of-trees model, and subsequently generate "Missing Not at Random (MNAR)" missingness in the response using a probit model. 

```{r}
set.seed(123456789)

n = 50
p = 2 #no. of dependent variables
q = 2 #no. of independent variables

x = matrix(runif(n*q, min = 0, max = 1), ncol = q)

sum_mu = matrix(0, ncol = p, nrow = n)
trees = 5
ome = matrix(c(1.7, 0.377, 0.377, 0.62), ncol=2)
kappa = 16*trees
true_trees = vector(mode = "list", length = trees)
for(i in 1:trees){
  df1 <- data.frame(matrix(ncol = 7, nrow = 1))
  colnames(df1) = c("parent", "lower", "upper", "split_variable", "split_value", "depth", "direction")
  df1[1,] <- c(0,0,1,0,1,0,0)
  n_splits = sample(seq(1,8), 1)
  mu = multi_rMVN(matrix(0, ncol=p, nrow = n_splits+1), kappa*diag(1,p)) 

  for(j in 1:n_splits){
    new_tree = propose_tree(df1, x, seq(ncol(x)), 20, 10, 2)
    df1 = new_tree$new_df
  }

  true_trees[[i]] = df1
  sum_mu = sum_mu + mu[new_tree$change_points,,drop=FALSE]
}

y = multi_rMVN(mean_mat = sum_mu, precision = ome)
y_original = y

plot(y_original[,1], y_original[,2])
```

### Simulate missing values 
```{r}
include_x = FALSE
include_y = TRUE
r = 1
if(include_x & !include_y){
  r = 1 + q
} else if (!include_x & include_y){
  r = 1 + p
} else if (include_x & include_y){
  r = 1 + p + q
}
corR = diag(1,p)

for(seed in sample(seq(1000,100000), size = 10000)){
  set.seed(seed)
  Psi_1 = matrix(rnorm(r*r), ncol=r) + diag(5, r) #rInvWishart(1, r+1, diag(1,r))[,,1]
  B = matrnorm(matrix(0, nrow=r, ncol=p), Psi_1, corR)
  if(include_x & !include_y){
    phi = cbind(rep(1, n), x) %*% B
  } else if (!include_x & include_y){
    phi = cbind(rep(1,n), y) %*% B
  } else if (include_x & include_y){
    phi = cbind(rep(1,n), x, y) %*% B
  } else {
    phi = matrix(B, nrow = n, ncol = p, byrow = TRUE)
  }
  z_mod = multi_rMVN(phi, solve(corR))
  m = matrix(1, nrow=n, ncol=p)
  m[z_mod<=0] = 0
  
  mis_prop = colSums(m)/n
  if(all(mis_prop > 0.9) & all(mis_prop < 0.99)) break
}

y[m==0] = NA

missing_index = which(is.na(y))
obs_index = which(!is.na(y))

```

Now that we have a dataset with MNAR missingness in the response, we can run `missBART.probit()`. 

```{r}
missBART = missBART.probit(x, y, burn = 100, iters = 100, thin = 1, n_trees = 10, include_x = FALSE, include_y = TRUE, scale = FALSE, show_progress = FALSE)

y_pred = missBART$y_pred
y_pred = Reduce("+", y_pred)/length(y_pred)

plot(y_pred[,1], y_original[,1])
plot(y_pred[,2], y_original[,2])
```
